%% Appendix: Collected Solicitations — Upwork Account Sharing Scheme

\documentclass[11pt]{article}

\usepackage[margin=1.2in]{geometry}
\usepackage{fontspec}           % LuaLaTeX/XeLaTeX
\setmainfont{TeX Gyre Pagella}
\usepackage{microtype}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{mdframed}
\usepackage{titlesec}
\usepackage{parskip}
\usepackage{enumitem}
\usepackage{fancyhdr}

\hypersetup{
  colorlinks=true,
  linkcolor=black,
  urlcolor=blue!60!black,
}

%% Header / footer
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small\textit{Appendix: Upwork Solicitations}}
\fancyhead[R]{\small\thepage}
\renewcommand{\headrulewidth}{0.4pt}

%% Message environment
\newmdenv[
  linewidth=0.4pt,
  linecolor=gray!60,
  backgroundcolor=gray!5,
  leftmargin=0pt,
  rightmargin=0pt,
  innerleftmargin=14pt,
  innerrightmargin=14pt,
  innertopmargin=10pt,
  innerbottommargin=10pt,
  skipabove=8pt,
  skipbelow=4pt,
]{msgbox}

\newcommand{\sender}[2]{%
  \medskip
  \noindent{\small\textbf{From:} #1 \hfill \textbf{Date:} #2}\par
  \smallskip
}

%% Section styling
\titleformat{\section}{\large\bfseries}{}{0em}{}[\titlerule]
\titleformat{\subsection}{\normalsize\bfseries\itshape}{}{0em}{}

%%-------------------------------------------------------------------
\begin{document}

\section*{Appendix~A: Collected Solicitations Offering Use of Third-Party
Upwork Accounts}

\bigskip

\noindent This appendix reproduces, verbatim and in chronological order, a
representative sample of unsolicited messages received via GitHub (and
associated contact channels) by the author. Each message proposes some variant
of the same scheme: a blocked or otherwise restricted Upwork account-holder
seeks access to a third party's legitimate Upwork account, offering revenue
sharing in exchange. The messages are presented without editorial alteration
beyond light formatting for legibility. Pseudonymous contact handles and
externally-supplied URLs have been retained as received.

\bigskip

%%-------------------------------------------------------------------
\subsection*{Message 1}

\sender{Roman}{2026-04-18}

\begin{msgbox}
Thank you for your reply.

I would like to discuss the possibility of collaboration with you. I have
primarily been active on the Upwork platform. Recently, I have been unable to
use my Upwork platform due to an issue. Therefore, if you possess or can create
Upwork platforms, we can collaborate to acquire projects together. I have
extensive technical experience with any task, so I can handle it. You can secure
the project, and I will take full responsibility for its execution.

I can also technically assist you in getting a job at a company. Once you get
hired, I will help you with all the work tasks. Alternatively, if you grant me
access to the platform via AnyDesk, I can manage the platform securely. Upon
acquiring and completing the project, I will give you 20–50\% of the earnings.

I am looking forward to hearing from you.

\smallskip
\noindent\textit{You can set the schedules here.}
\url{https://calendar.app.google/wgk8Zw1ZExEJKAR76}\\
\textit{You can also reach me on Telegram:} \texttt{@jareddmartin}
\end{msgbox}

\newpage
%%-------------------------------------------------------------------
\subsection*{Message 2}

\sender{Kratos}{2026-03-17}

\begin{msgbox}
I am a fullstack developer from Singapore and I have 9+ years of experience in
full stack web3 and AI engineering. I have worked for local companies as a
full-stack developer so far but now I want to work as a freelancer on Upwork
for more income. But unfortunately my account was blocked 3 years ago and an
Asian hourly rate is usually lower than a European or American. So I am looking
for someone to help me to work on Upwork. I will pay 10 percent of earnings on
Upwork per month and have a long term relationship with you.

\smallskip\noindent\textbf{Why do I need to use the European Upwork account?}

As you know, Europeans are easier than Asians to work with a high hourly rate
on Upwork. Honestly, it is so difficult to get good jobs using the Asian Upwork
account on Upwork. So I don't want to use a Chinese (Asian) account and only
use the European Upwork account.

\smallskip\noindent\textbf{How can I use your Upwork account?}

If you have an Upwork account and don't use it, I wanna use it on your local
computer via Teamviewer or Anydesk. If you haven't got it, you can create a new
account on Upwork. Also, I think if you use your account part-time, you can use
it with you. Only I will use your account while you don't use it.

\smallskip\noindent\textbf{How can we transfer money with each other?}

You can add Payoneer or PayPal to your Upwork account and if I earn some money
using your account, you can withdraw it via your Payoneer or PayPal. And then
you can get 10\% of the money from there and you can send the rest of the money
to my PayPal or Payoneer.

\smallskip\noindent\textbf{Why do I need access to your local via Teamviewer or
AnyDesk?}

If I use your Upwork account using a virtual server such as VPN, VPS, AWS,
Upwork detects the proxy server IP and blocks your account. So if I can use
your account, I must absolutely use it on your local computer via Teamviewer or
AnyDesk. If you worry about your personal secret data on your computer, you can
set a VM on your computer and I can use your account on the VM of your computer.

\newpage

\smallskip\noindent\textbf{You don't have to worry about me stealing your Upwork
account. Why?}

Honestly there isn't a need to have your Upwork account without you. If I use
your account on another location using VPN, VPS, your account will be blocked
because of an IP location change. Also, Upwork requires video chats sometimes
and I really can't do it without your help.

\smallskip
I found you on Github. I look forward to hearing from you soon.
\end{msgbox}

%%-------------------------------------------------------------------
\subsection*{Message 3}

\sender{Evan North [Noriel]}{2025-10-09}

\begin{msgbox}
Hello, My name is Noriel, a software engineer with over 8 years of experience
in Python, AWS, and full stack development. And I also have proven experience in
Upwork for over 3 years, and during that time, I've got a good monthly income,
just around \$8{,}000–\$10{,}000, but recently my account has been blocked with
the problem related to my critical fault. It was about outside contact without
Upwork. The reason was that I and my clients wanted to make a long term
relationship and also for the convenience of the communication. But anyway, this
is all related to violation activities according to Upwork's policy. So,
unfortunately my account was blocked last month. Because of this, I can't
continue my freelance work anymore.

But if you don't mind, we can collaborate together. So, I'm going to use your
Upwork with your name with my experiences, skills and stack. And after getting a
good task on Upwork, I will handle that task, and after that, I will share your
40\% of monthly income for using your Upwork. It will be
\$2{,}000–\$4{,}000 or more according to our effort.

And, to do this work, we need to create Upwork and pass the verification.

\smallskip
\noindent\textit{Contact:} Telegram: \texttt{@bestdiamond9},
WhatsApp: \texttt{+63 936 935 7985}
\end{msgbox}


\newpage
%%-------------------------------------------------------------------
\subsection*{Message 4}

\sender{Douglas}{2026-02-25}

\begin{msgbox}
Hi, thanks for your interest in my proposal.

I hope you may have any job site (e.g.\ Upwork) accounts which are inactive.
So what I am proposing is that instead of keeping inactive, you can allow me to
bid on jobs through your account. I can handle bidding on jobs, completing the
projects, and delivering the final results through your accounts---essentially
managing the full end-to-end workflow. Your involvement would be minimal, so
you can earn passive income without needing to participate actively.

If possible, I'd like access to your account so I can place bids directly. If
that isn't workable, we can use a second option where I send you job links daily
and you place the bids manually---though this may become inconvenient for you.
I will cover the cost of purchasing connections for bidding. For account access,
I would use a separate device or a virtual machine to ensure your personal data
remains completely isolated and inaccessible.

\smallskip\noindent\textbf{Regarding income sharing:}
\begin{itemize}[nosep, leftmargin=1.5em]
  \item For standard projects: 15\,/\,85 split (you take 15\%).
  \item For projects requiring calls from your side---initially only: 20\,/\,80
  split (you take 20\%).
\end{itemize}

I've successfully done this before with another collaborator, and we earned
about \$15{,}000 per month through Upwork.
\end{msgbox}

\newpage
%%-------------------------------------------------------------------
\subsection*{Message 5}

\sender{Tonny}{2025-11-19}

\begin{msgbox}
Hello, bro. Nice to meet you. I am a full stack and blockchain developer. I am
looking for a partner to support my account like LinkedIn and Upwork. Let's
discuss more details via X.

\smallskip
\noindent\texttt{@0xFortuneRust} — Please send me messages.
\end{msgbox}

%%-------------------------------------------------------------------
\subsection*{Message 6}

\sender{Carter Grant}{2025-11-03}

\begin{msgbox}
Hello. I am Carter Grant and a web developer in Hong Kong. I recently came
across your profile on Github and was really impressed with your education
background and skills.

Heretofore, I have worked as a full-stack developer in several freelancing
platforms, but it is really hard to take jobs as a freelancer so I have decided
to collaborate with a senior developer because my job interview skills are poor.
Every time I tried the interview, I failed because of my verbal English and
Asian accent. But I can perfectly understand English and I'm really good at
coding. Also I am a very professional worker, so I can handle at least 2 or
more full-time jobs myself.

So I think you can focus on getting more jobs and we will be able to take more
remote jobs. For the compensation, we can share the budget from the jobs.

My main skills are React, Next.js, Vue, Angular, Python, Node, LLM, RAG and
nearly everything related to JavaScript and TypeScript.

These are my previous works: \url{https://siteauditpro.com} and
\url{https://esmartcontrol.com}.

You may have no project or you can't forward your current work for some reasons,
then we can try to take jobs from Upwork. Actually in Upwork, the most boring
and time-taking process is to send proposals. I can handle the proposal
submissions, and if we land a job, I'll take care of the development work and
you'd just need to have an interview with the client when it is necessary.
\end{msgbox}

\newpage
%%-------------------------------------------------------------------
\subsection*{Message 7}

\sender{Richard}{2025-10-19}

\begin{msgbox}
Hello Cogito. I'm an Upwork agency. I manage my own account and other people's
accounts. I'd like to lease your Upwork account to me. I'll pay you \$200 per
month for two months, and then 20\% of the Upwork account's earnings after
3 months. You just need to create an account and verify. If you like my offer,
please contact me at the contact information below. I won't lie. If you're
interested in this project, it's entirely your choice.

\smallskip
\noindent WhatsApp: \texttt{+1 (540) 831 8619},
Telegram: \texttt{@Ricahrd\_PM33}
\end{msgbox}

%%-------------------------------------------------------------------
\subsection*{Message 8}

\sender{Ninja Dev}{2025-07-22}

\begin{msgbox}
Hi, I hope you are lucky in everything you do. I am from Japan. I've been
freelancing on Upwork for over 1.5 years and have racked up over \$100{,}000.
Most of my clients are in the U.S., but the time zone shuffle has caused a few
hiccups, and I've lost some clients along the way.

So, I thought---why not team up? If you create an account in your area, I can
take care of client communication, proposals, and project management. We can
share the profits, and I bet we can work smarter, not harder!

\smallskip
\noindent Telegram: \texttt{wm\_0411}, Discord: \texttt{together\_withu}
\end{msgbox}

\newpage
%%-------------------------------------------------------------------
\subsection*{Message 9}

\sender{Bohdan}{2025-07-12}

\begin{msgbox}
Nice to meet you---hope you're doing well! I came across you on Github
(\url{https://github.com/standardgalactic}).

I'm Bohdan, a software engineer from Ukraine with 8+ years of experience,
especially in full-stack development and AI-driven applications. I've been
freelancing for several years, primarily through Upwork, and have built a solid
client base and track record there.

\smallskip\noindent\textbf{Why I'm reaching out:} Due to recent account
restrictions, I'm currently unable to operate my own Upwork account---which is
why I'm looking for a trusted partner who can help me continue doing what I do
best.

\smallskip\noindent\textbf{What the collaboration would look like:}
\begin{itemize}[nosep, leftmargin=1.5em]
  \item You create (or use an existing) Upwork account.
  \item I manage everything---bidding, proposals, client communication,
  delivering the work.
  \item You receive the funds (they go to your account).
  \item We split the profits fairly.
  \item You can be as hands-off or involved as you prefer.
\end{itemize}

The goal is to create a reliable income stream for both of us, with minimal
effort on your side and no risk---since I only get paid after the job is
completed and paid for. I understand this kind of partnership is based on trust,
so I'm more than happy to share my portfolio, hop on a quick call, or start
small and build step-by-step.
\end{msgbox}

\bigskip

\noindent\rule{\linewidth}{0.4pt}

\medskip
\noindent\textit{End of Appendix~A.}

\end{document}
